---
title: "Development of a reservoir morphology dataset to inform modeling greenhouse gas emissions from U.S. reservoirs."
author: "Jeffrey Hollister, Jake Beaulieu, Bridget Deemer, and Alex Hall"
format: pptx
editor: visual
editor_options: 
  chunk_output_type: console
---

```{r, echo=FALSE}
source(here::here("R/packages.R"))
source(here("R/functions.R"))
```

## Introduction

Understanding ecological processes in lentic systems, such as reservoirs, often requires the quantification of metrics of the waterbody's shape and size (i.e. morphology). In support of ongoing efforts in the United States (U.S.) to better characterize greenhouse gas emissions from reservoirs, we have expanded past efforts (e.g. National Hydrography Dataset Plus), and are building a database of lake morphology for U.S. reservoirs.

## Data Sources

We use publicly available datasets to first build a database of U.S. reservoirs and then calculate the metrics for each using open source geospatial software (e.g. R). Existing databases with relevant reservoir morphology metrics, such as LAGOS and the National Inventory of Dams, are also included. The database includes estimates for 147 reservoirs measured as part of the Survey of Greenhouse Gas Emissions project led by the U.S. Environmental Protection Agency and efforts are underway to expand the number of included reservoirs.

## Database Development

We currently calculate 13 lake morphometry metrics including descriptors of lake shape (e.g. area, shoreline length), depth (e.g. mean depth, volume), and fetch. We present our procedures for developing the reservoir database and approaches for calculating each of the metrics.

### The Data

-   SuRGE Polygons
    -   Two separate efforts
        -   Original effort in XXXX
            -   n =
            -   n sites per reservoir =
                -   depth at each taken from existing bathymetry
        -   SuRGE in 2020-2023
            -   n =
            -   n sites per reservoir \~ 15
                -   depth measured at each
-   Existing Datasets
    -   SuRGE Crosswalk to
        -   List
-   Output geopackage
    -   Metrics for each resrvoir
    -   Point dataset with measured depth
        -   List sources

            ```         
            "surge sites"  "nla 12 phab"  "nla 17 phab" "nla 22 phab"  "nla07 index" "nla17 index" "nla12 index" "nla22 index"  "pre-SuRGE resevoris sampling loaction - bathymetry" "pre-SuRGE reservoir bathymetry max depth"    
            ```
-   Output flat file
-   These reservoirs had an average size of 516 hectares, predicted mean depth on average was 10 meters, and average maximum fetch was 2762 meters.
-   The final database will be publicly available and will support modeling of greenhouse gas emissions for the U.S. Inventory of Greenhouse Gas Emissions and Sinks.

## Next Steps

-   Best available data
    -   many sources of maximum depth
-   Volume calculations
-   Additional Sources?

## Links and References

-   GitHub Repo:
-   Lakemorpho
-   data source links

## Figures

```{r include=FALSE}
surge_poly <- st_read(here("data/surge/all_lakes.gpkg"), layer = "all_lakes")  |>
  st_transform(5072)
surge_poly <- mutate(surge_poly, 
                     study = case_when(as.numeric(lake_id) >= 1000 ~
                                         "2016 Study",
                                       TRUE ~ "SuRGE"))
surge_poly <- filter(surge_poly, as.numeric(lake_id) != 1033)
us <- USAboundaries::us_states() |>
  filter(!state_abbr %in% c("AK", "HI")) |>
  st_transform(5072)
nat_map <- surge_national_map(surge_poly, us)
ggsave(here("products/hollister_et_al_agu_present/figures/nat_map.jpg"), nat_map, width = 24, height = 18, units = "in", dpi = 600)
```

```{r include = FALSE}
surge_all_pts <- st_read(here("data/surge_morpho_point_depth.gpkg"))  |>
  st_transform(5072)
lake_1013 <- filter(surge_poly, lake_id == 1013)
lake_pts_1013 <- filter(surge_all_pts, lake_id == 1013)
res_map <- single_reservoir_map(lake_1013, lake_pts_1013)
ggsave(here("products/hollister_et_al_agu_present/figures/res_map.jpg"), width = 18, height = 18, units = "in", dpi = 600)
```
